---
title: "Lab 1"
author: "Becca Reynolds"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Attach packages
```{r}
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
```
Read in NOAA Fisheries data
Use here() if .csv is not in your working directory. You can also use "data/noaa_fisheries.csv" but it's safer to use here() because it can ALSO pull from upstream.
```{r}
us_landings <- read_csv(here("data","noaa_fisheries.csv"))
```
Tidy data
```{r}
landings_tidy <- us_landings %>% 
  janitor::clean_names() %>% 
  mutate(state = str_to_lower(state),
         afs_name = str_to_lower(afs_name)) %>% 
  mutate(dollars_num = parse_number(dollars_usd))
```
#^^dollars_usd is going to be read as a character because you have at least one non-numeric thing ($) in the observation. We want to get rid of the ($) and store as numeric
There are many different functions in the readr package to parse things, like ?parse_number
Use this this with mutate on dollars_usd to drop non-numeric characters


Get info for salmon:
```{r}
salmon_landings <- landings_tidy %>% 
  mutate(afs_clean = str_remove(afs_name, pattern = "aggregate")) %>% 
  filter(str_detect(afs_clean, pattern = "salmon")) %>% 
  separate(afs_clean, into = c("group", "species"), sep = ",")

```
^^you can filter by looking for a partial match instead of an exact match, keep any row with the pattern shark in it. filter by asking: is this string detected? if true, keep that row, if false, don't keep that row. After afs_clean, into = c("group", "species"), tell R to seperate at the comma after salmon in afs_clean observations so it knows how to seperate strings in the column of interest.

Find some grouped summary data:

Find annual total of US landings and dollar value by summing across all states for each type of salmon using group_by() and summarize()

SO group by year and species...
```{r}
salmon_summary <- salmon_landings %>%
  group_by(year, species) %>% 
  summarize(
    tot_landings = sum(landings_pounds),
    tot_value = sum(dollars_num)
  )
```

Make a graph to visualize this
```{r}
salmon_landings_graph <- ggplot(data = salmon_summary, 
                                aes(x = year, y = tot_landings)) +
  geom_line(aes(color = species)) +
  theme_minimal()
  
salmon_landings_graph

```
You could also group by species like this:
ggplot(data = salmon_summary, 
                                aes(x = year, y = tot_landings, group = species)) +
  geom_line()
  
  But it won't assign colors to each species like it does using the code in the code chunk above.

Export graph
```{r}
ggsave(plot = salmon_landings_graph,
       here("figures", "us_salmon_br.png"),
       height =5,
       width = 8)
```
save the graph to the foler "figures" in this repo. here let's you do this. you can specify dpi, size, ...etc.

Make a table
```{r}
salmon_first_5 <- salmon_summary %>% 
  head(5)
```

```{r}
kable(salmon_first_5) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)
```

